Reproducible development environments with rix

### 2023-12-18 paper.qmd
some text b b

[Nix]{.proglang} is a package manager that can be used to install and build
software in a completely reproducible manner. With [Nix]{.proglang}
it is possible to install [R]{.proglang} and [R]{.proglang} packages as well
(whether these come from CRAN or Bioconductor)

is that its repository of software is huge. As of writing, it contains more than
80.000 packages, and the entirety of CRAN and Bioconductor is available through
Nix’s repositories. This means that using Nix, it is possible to install not
only R, but also all the packages required for your project. The obvious
question is why use Nix instead of simply installing R and R packages as usual.
The answer is that Nix makes sure to install every dependency of any package, up
to required system libraries. For example, the `{xlsx}` package requires the
Java programming language to be installed on your computer to successfully
install. This can be difficult to achieve, and `{xlsx}` bullied many R
developers throughout the years (especially those using a Linux distribution,
`sudo R CMD javareconf` still plagues my nightmares). But with Nix, it suffices
to declare that we want the `{xlsx}` package for our project, and Nix figures
out automatically that Java is required and installs and configures it. It all
just happens without any required intervention from the user. The second
advantage of Nix is that it is possible to pin a certain revision of the Nix
packages’ repository (called nixpkgs) for our project. Pinning a revision
ensures that every package that Nix installs will always be at exactly the same
versions, regardless of when in the future the packages get installed.

>>>>

[Nix]{.proglang} is a package manager that can be used to install and build
software in a completely reproducible manner. As of writing, it contains more
than 80.000 packages, and the entirety of CRAN and Bioconductor is available
through [Nix]{.proglang}’s repositories. This means that using [Nix]{.proglang},
it is possible to install not only [R]{.proglang}, but also all the packages
required for a project. The reason why one should use [Nix]{.proglang} to
install [R]{.proglang} packages, and not use the usual, built-in,
`install.packages()` function instead, is that when installing a package with
[Nix]{.proglang}, [Nix]{.proglang} makes sure to install every dependency of
every package, whether this dependency is another R package or a system-level
dependency.

For example, the `{xlsx}` package requires the Java programming language to be
installed on your computer to successfully install. This can be difficult to
achieve, and `{xlsx}` bullied many R developers throughout the years (especially
those using a Linux distribution, `sudo R CMD javareconf` still plagues my
nightmares). But with Nix, it suffices to declare that we want the `{xlsx}`
package for our project, and Nix figures out automatically that Java is required
and installs and configures it. It all just happens without any required
intervention from the user. The second advantage of Nix is that it is possible
to pin a certain revision of the Nix packages’ repository (called nixpkgs) for
our project. Pinning a revision ensures that every package that Nix installs
will always be at exactly the same versions, regardless of when in the future
the packages get installed.
